---
title: 'FIV CourseWork'
subtitle: 'IBM Employee Attrition Survey EDA' 
author: '20299113'
classoption: twoside
output:
  html_notebook:
    toc: no
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document:
    toc: no
    fig_caption: yes
    number_sections: true
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
header-includes: 
- \usepackage{graphicx}
- \usepackage{float}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[CE,CO]{}
- \fancyhead[LE,LO]{\textit{20299113}}
- \fancyhead[RE,RO]{\nouppercase{\textit{\leftmark}}}
- \usepackage{xcolor}
- \usepackage{framed}
- \colorlet{shadecolor}{gray!10}
---
# 0. Data Description
## 0.1. DataSource
[IBM HR Analytics Employee Attrition & Performance](https://www.kaggle.com/pavansubhasht/ibm-hr-analytics-attrition-dataset)

## 0.2. Inspiration/Reference
1. https://www.kaggle.com/ysjang0926/analysis-of-employee-attrition-eda-and-modeling
2. https://www.kaggle.com/yafeifionajian/ibm-attrition-visualization
3. https://www.kaggle.com/pelinpektekin/eda-and-data-visualization-for-each-var
4. https://www.kaggle.com/gsdeepakkumar/whose-turn-over-it-is

## 0.3. Features
* Age: The age of employee. Numerical
* Attrition: The state of employment. Logical/Categorical: Yes/No
* BusinessTravel: The frequency of business travel. Categorical. 
* DailyRate: Numerical
* Department: The belonging department. Categorical.
* DistanceFromHome: The distance between home and company. Numerical
* Education: The education level. Numerical/Categorical. (1 'Below College',2 'College',3 'Bachelor',4 'Master',5 'Doctor')
* EducationField: The education field. Categorical.
* EmployeeCount: Numerical.
* EmployeeNumber: The Employee ID. Numerical.
* EnvironmentSatisfaction: The satisfaction score from employee. Numerical.(1 'Low',2 'Medium',3 'High',4 'Very High')
* Gender: Categorical: Male/Female.
* HourlyRate: Numerical.
* JobInvolvement: Numerical/Categorical. (1 'Low',2 'Medium',3 'High',4 'Very High')
* JobLevel: Numerical/Categorical.        
* JobRole: The job title. Categorical. 
* JobSatisfaction: Numerical/Categorical. (1 'Low',2 'Medium',3 'High',4 'Very High')
* MaritalStatus: The marital status. Categorical.
* MonthlyIncome: Numerical.
* MonthlyRate:Numerical.      
* NumCompaniesWorked: The number of companies used to work. Numerical.
* Over18: Is the age over 18? Categorical: Y/N.
* OverTime: Work overtime. Logical/Categorical: Yes/No
* PercentSalaryHike: The hike% of salary. Numerical.
* PerformanceRating: Categorical. (1 'Low',2 'Good',3 'Excellent',4 'Outstanding')
* RelationshipSatisfaction: The Relationship Satisfaction Level. Numerical/Categorical. (1 'Low',2 'Medium',3 'High',4 'Very High')
* StandardHours: Numerical.
* StockOptionLevel: Categorical.
* TotalWorkingYears: Total working years. Numerical.
* TrainingTimesLastYear: Training times during last year.
* WorkLifeBalance: Work&Life balance level. Numerical/Categorical.(1 'Bad',2 'Good',3 'Better',4 'Best')
* YearsAtCompany: Total working years at current company. Numerical.
* YearsInCurrentRole: Total working years at current role. Numerical.
* YearsSinceLastPromotion: The duration(year) since last promotion. Numerical.
* YearsWithCurrManager: The total years with the current manager. Numerical.


# 1. Data Preprocessing
## 1.1. Load packages
```{r}
# Load related packages quietly
library(mice, quietly = TRUE)
library(psych, quietly = TRUE)
library(ggplot2, quietly = TRUE)
library(magrittr, quietly = TRUE)
library(plotly, quietly = TRUE)
library(gridExtra, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(GGally, quietly = TRUE)
library(data.table, quietly = TRUE)
```

## 1.2. Load the DataSet
```{r}
dataSet = read.csv('WA_Fn-UseC_-HR-Employee-Attrition.csv')
print(paste("Rows:",dim(dataSet)[1],", Columns:",dim(dataSet)[2]))
# Show the first 5 rows
head(dataSet)
```

## 1.3. DataSet Check
```{r}
md.pattern(dataSet)
```

## 1.4. Data Type Check
```{r}
str(dataSet)
```

## 1.5. Show the basis statistics
```{r}
describe(dataSet)
```

## 1.6. Feature Virtualization
```{r}
## Interactive Graph
# https://plotly.com/r/sunburst-charts/
fig <- plot_ly(
  labels = c("Features", "Background", "Position", "Satisfaction", "Performance", "Payment", "WorkingENV",
             "Gender", "Age", "Education", "EducationField", "MaritalStatus", "TotalWorkingYears", "NumCompaniesWorked", "Attrition",
             "HourlyRate", "DailyRate", "MonthlyRate", "MonthlyIncome", "PercentSalaryHike", "StockOptionLevel",
             "JobRole", "JobLevel", "Department", "YearsWithCurrManager", "YearsInCurrentRole", "YearsAtCompany",
             "DistanceFromHome", "StandardHours", "OverTime", "BusinessTravel", "TrainingTimesLastYear", "YearsSinceLastPromotion",
             "JobSatisfaction","RelationshipSatisfaction","EnvironmentSatisfaction","WorkLifeBalance",
             "JobInvolvement", "PerformanceRating"),
  parents = c("", "Features", "Features", "Features", "Features", "Features", "Features",
              "Background", "Background", "Background", "Background", "Background", "Background","Background","Background",
              "Payment", "Payment", "Payment", "Payment", "Payment", "Payment",
              "Position","Position","Position","Position","Position","Position",
              "WorkingENV","WorkingENV","WorkingENV","WorkingENV","WorkingENV","WorkingENV",
              "Satisfaction","Satisfaction","Satisfaction","Satisfaction",
              "Performance", "Performance"),
  type = 'sunburst'
)
fig %>% layout(title = '<b>Feature Visualization</b>',
               margin = list(t=60),
               font = list(size=12))
```

# 2. Initial questions
## 2.1. What is the distribution of the employee attrition?
```{r}
## ggplot2
# Create an attrition dataframe 
df_attrition = transform(as.data.frame(table(dataSet$Attrition)),percentage=Freq/nrow(dataSet)*100, Attrition = c("Non-Resignation","Resignation"))

# Draw a bar graph 
p_attrition = ggplot(data = df_attrition[c("Freq","Attrition")], aes(x = '', y = Freq, fill = Attrition))+
  geom_bar(stat = 'identity', position = 'stack')

# Transform the bar graph to pie graph
p_attrition + 
  coord_polar(theta = 'y')+
  theme_void()+
  ggtitle("Attrition distribution of employee in IBM Survey")+
  geom_text(aes(label = paste(round(df_attrition$percentage, 1), "%","(",df_attrition$Freq,")", sep = "")), position = position_stack(vjust = 0.55), color = "white")+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 16), legend.position="right")
```

```{r}
##  Interactive Graph
# https://plotly.com/r/pie-charts/

# Choose a colour set
colors = c('rgb(118,218,145)','rgb(248,149,136)')

fig_attrition = plot_ly(df_attrition, labels = ~Attrition, values = ~Freq, type = 'pie', marker = list(colors = colors))
fig_attrition = fig_attrition %>% layout(title = '<b>Attrition distribution of employee in IBM Survey</b>',
                                         margin = list(t=60),
                                         font = list(size=16),
                                         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                                         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                                         legend=list(title=list(text='<b> Attrition </b>')))
fig_attrition
```

## 2.2. How does age influence employee attrition?
```{r}
## ggplot2
p_age = ggplot(dataSet, aes(x=Age, fill=Attrition, color=Attrition))
p_age+
  geom_density(position="identity", alpha=0.6)+ 
  theme_light()+ 
  scale_fill_brewer(palette="Pastel1")+
  ggtitle("Age & Employee Attrition")+ 
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 16), legend.position="top")+
  labs(x = "Age") 
```

```{r}
## Interactive Graph
# https://plotly.com/r/filled-area-plots/
data_yes = dataSet[which(dataSet$Attrition == "Yes"),]
density1 = density(data_yes$Age)

data_no = dataSet[which(dataSet$Attrition == "No"),]
density2 = density(data_no$Age)
","
fig_age = plot_ly(x = ~density1$x, y = ~density1$y, type = 'scatter', mode = 'lines', name = 'Resignation', fill = 'tozeroy')
fig_age = fig_age %>% add_trace(x = ~density2$x, y = ~density2$y, name = 'Non-Resignation', fill = 'tozeroy')
fig_age = fig_age %>% layout(title = '<b>Age & Employee Attrition</b>',
                             margin = list(t=60),
                             font = list(size=16),
                             xaxis = list(title = 'Age'),
                             yaxis = list(title = 'Density'))
fig_age
```

## 2.3. How does JobLevel influence employee attrition?
```{r}
## ggplot2
p = ggplot(dataSet, aes(x=Attrition, group=JobLevel))
p +
  geom_bar(aes(y=..prop.., fill=factor(..x..)), stat="count",show.legend=FALSE)+
  theme_bw()+
  facet_grid(~factor(JobLevel))+
  scale_y_continuous(labels=scales::percent)+
  ggtitle("JobLevel & Employee Attrition")+
  geom_text(aes(label = scales::percent(..prop..,accuracy = 0.1), y = ..prop.. ),stat= "count",vjust =-0.2, size=3.5)+
  theme(plot.title = element_text(hjust = 0.5, size=14, face="bold"))+
  labs(x = "Attrition", y = "Percentage")
```

## 2.4. How does OverTime influence employee attrition?
```{r}
## ggplot2
attrition_y = dataSet$OverTime[dataSet$Attrition == "Yes"]
attrition_n = dataSet$OverTime[dataSet$Attrition == "No"]

num_attrition_y = length(attrition_y)
num_attrition_n = length(attrition_n)

df_attrition_y = transform(as.data.frame(table(attrition_y)),percentage=Freq/num_attrition_y*100, Overtime = c("No","Yes"))
df_attrition_n = transform(as.data.frame(table(attrition_n)),percentage=Freq/num_attrition_n*100, Overtime = c("No","Yes"))

# Draw a bar graph 
p1 = ggplot(data = df_attrition_y[c("Freq","Overtime")], aes(x = '', y = Freq, fill = Overtime))+
  geom_bar(stat = 'identity', position = 'stack')

p2 = ggplot(data = df_attrition_n[c("Freq","Overtime")], aes(x = '', y = Freq, fill = Overtime))+
  geom_bar(stat = 'identity', position = 'stack')

# Transform the bar graph to pie graph
p_attrition_y = p1+ 
  coord_polar(theta = 'y')+
  theme_void()+
  ggtitle("Resignation")+
  geom_text(aes(label = paste(round(df_attrition_y$percentage, 1), "%","(",df_attrition_y$Freq,")", sep = "")), position = position_stack(vjust = 0.5), color = "white")+
  theme(plot.title = element_text(hjust = 0.5, size = 12))

p_attrition_n = p2+ 
  coord_polar(theta = 'y')+
  theme_void()+
  ggtitle("Non-Resignation")+
  geom_text(aes(label = paste(round(df_attrition_n$percentage, 1), "%","(",df_attrition_n$Freq,")", sep = "")), position = position_stack(vjust = 0.5), color = "white")+
  theme(plot.title = element_text(hjust = 0.5, size = 12))

# Show two plots in graph
# https://rpkgs.datanovia.com/ggpubr/reference/annotate_figure.html
figure = ggarrange(p_attrition_y, p_attrition_n, nrow=1, common.legend = TRUE, legend="bottom")
annotate_figure(figure, top = text_grob("OverTime distribution on Attrition", face = "bold", size = 14))
```

```{r}
##  Interactive Graph
# https://plotly.com/r/pie-charts/

# Choose a colour set
colors = c('rgb(118,218,145)','rgb(248,149,136)')

# Add annotations for each sub pie chart
annotations=list(
    list(
      x = 0.225, 
      y = 0.95, 
      font = list(size = 15), 
      text = "Resignation", 
      xanchor = "center", 
      yanchor = "bottom", 
      showarrow = FALSE),
    
    list(
      x = 0.775, 
      y = 0.95, 
      font = list(size = 15), 
      text = "Non-Resignation", 
      xanchor = "center", 
      yanchor = "bottom", 
      showarrow = FALSE))

# Show two plots in graph
fig = plot_ly()
fig = fig %>% add_pie(df_attrition_y, labels = ~df_attrition_y$Overtime, values = ~df_attrition_y$Freq, type = 'pie', marker = list(colors = colors), domain = list(row = 0, column = 0))
fig = fig %>% add_pie(df_attrition_n, labels = ~df_attrition_n$Overtime, values = ~df_attrition_n$Freq, type = 'pie', marker = list(colors = colors), domain = list(row = 0, column = 1))
fig = fig %>% layout(annotations = annotations)
fig = fig %>% layout(title = '<b> OverTime distribution on Attrition </b>',
                     margin = list(t=60),
                     font = list(size=16),
                     grid=list(rows=1, columns=2),
                     xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                     yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                     legend=list(title=list(text='<b> Overtime </b>'), xanchor = "top", yanchor="top"))

fig
```

## 2.5. How does Job Satisfaction influence employee attrition?
```{r}
# ggplot2
p_job_sat = ggplot(dataSet, aes(x=Attrition, y=JobSatisfaction, color=Attrition, fill=Attrition))
p_job_sat+ 
  geom_boxplot()+
  theme_light()+
  ggtitle("Job Satisfaction & Employee Attrition")+   
  scale_fill_brewer(palette="Pastel1")+
  stat_summary(fun=mean, geom="crossbar", color="black", width = 0.75, size= 0.2, fatten = 1.0)+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 16))+
  labs(x = "Attrition", y = "Job Satisfaction")
```
```{r}
## Interactive Graph
# https://plotly.com/r/box-plots/
data_yes = dataSet[which(dataSet$Attrition == "Yes"),]
box1 = data_yes$JobSatisfaction

data_no = dataSet[which(dataSet$Attrition == "No"),]
box2 = data_no$JobSatisfaction

fig <- plot_ly(y = ~box1, type = "box",name = 'Resignation', boxmean = T)
fig <- fig %>% add_trace(y = ~box2,name = 'Non-Resignation')
fig = fig %>% layout(title = "<b>Job Satisfaction & Employee Attrition</b>", yaxis=list(title = 'Job Satisfaction'),
                     margin = list(t=60),
                     font = list(size=16))
fig
```

## 2.6. How does WorkLifeBalance influence employee attrition?
```{r}
## ggplot2
# Transform numerical data to categorical data respectively
dataSet$WorkLifeBalance[dataSet$WorkLifeBalance == 1] = "Bad"
dataSet$WorkLifeBalance[dataSet$WorkLifeBalance == 2] = "Good"
dataSet$WorkLifeBalance[dataSet$WorkLifeBalance == 3] = "Better"
dataSet$WorkLifeBalance[dataSet$WorkLifeBalance == 4] = "Best"

# https://ggplot2.tidyverse.org/reference/scale_continuous.html
p = ggplot(dataSet, aes(x=Attrition, group=WorkLifeBalance))
p +
  geom_bar(aes(y=..prop.., fill=factor(..x..)), stat="count",show.legend=FALSE)+
  theme_bw()+
  facet_grid(~factor(WorkLifeBalance,levels = c("Bad","Good","Better","Best")))+
  scale_y_continuous(labels=scales::percent)+
  ggtitle("WorkLifeBalance & Employee Attrition")+
  geom_text(aes(label = scales::percent(..prop..,accuracy = 0.1), y = ..prop.. ),stat= "count",vjust =-0.2, size=3.5)+
  theme(plot.title = element_text(hjust = 0.5, size=14, face="bold"))+
  labs(x = "Attrition", y = "Percentage")
```

## 2.7. How does salary influence employee attrition?
```{r}
## ggplot2
p_monthly_income = ggplot(dataSet, aes(x=Attrition, y=MonthlyIncome, color=Attrition, fill=Attrition))
p_monthly_income+ 
  geom_boxplot()+
  theme_light()+
  ggtitle("Monthly Income & Employee Attrition")+   
  scale_fill_brewer(palette="Pastel1")+
  stat_summary(fun=mean, geom="crossbar", color="black", width = 0.75, size= 0.2, fatten = 1.0)+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 16))+
  labs(x = "Attrition", y = "Monthly Income ($)")
```
```{r}
## Interactive Graph
# https://plotly.com/r/box-plots/
data_yes = dataSet[which(dataSet$Attrition == "Yes"),]
box1 = data_yes$MonthlyIncome

data_no = dataSet[which(dataSet$Attrition == "No"),]
box2 = data_no$MonthlyIncome
fig = plot_ly(y = ~box1, type = "box",name = 'Resignation', boxmean = T)
fig = fig %>% add_trace(y = ~box2,name = 'Non-Resignation')
fig = fig %>% layout(title = "<b>Monthly Income & Employee Attrition</b>", yaxis=list(title = 'Monthly Income($)'),
                     margin = list(t=60),
                     font = list(size=16))

fig
```

## 2.8. Which JobRole earn the most Salary overall?
```{r}
## ggplot2
# https://ggplot2.tidyverse.org/reference/stat_summary.html
p = ggplot(dataSet, aes(x=reorder(JobRole,-MonthlyIncome), y=MonthlyIncome, color=JobRole, fill=MonthlyIncome))
p+
  geom_boxplot()+
  theme_light()+
  ggtitle("MonthlyIncome & JobRole")+   
  scale_fill_brewer(palette="Set1")+
  stat_summary(fun=mean, geom="crossbar", color="black", width = 0.75, size= 0.2, fatten = 1.5)+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 16),axis.text.x = element_text(angle = -45, hjust = 0, vjust = 0))+
  labs(x = "JobRole", y = "MonthlyIncome($)")
```
```{r}
## Interactive Graph
# https://plotly.com/r/box-plots/
ordered = c("Manager", "Research Director", "Healthcare Representative", "Manufacturing Director", "Sales Executive", "Human Resources", "Research Scientist", "Laboratory Technician", "Sales Representative")

fig = plot_ly(y = ~dataSet$MonthlyIncome, color = ~factor(dataSet$JobRole,levels = ordered), type = "box", boxmean = T)
fig = fig %>% layout(title = "<b>MonthlyIncome & JobRole</b>", yaxis=list(title = 'MonthlyIncome($)'),
                     margin = list(t=60))
fig
```


## 2.9. What's the relationship between TotalWorkingYears and MonthlyIncome ?
```{r}
## ggplot2
p = ggplot(dataSet, aes(x = TotalWorkingYears, y = MonthlyIncome))
p + geom_point(color = "steelblue")+
  ggtitle("TotalWorkingYears & MonthlyIncome")+
  theme(plot.title = element_text(hjust = 0.5, size=14, face="bold"))+
  labs(y = "Monthly Income($)")+
  geom_smooth(colour = "black", alpha=0.5, size=0.5)
```

```{r}
##  Interactive Graph
p = ggplot(dataSet, aes(x = TotalWorkingYears, y = MonthlyIncome))
p = p + geom_point(color = "steelblue")+
  ggtitle("TotalWorkingYears & MonthlyIncome")+
  theme(plot.title = element_text(hjust = 0.5, size=14, face="bold"))+
  labs(y = "Monthly Income($)")+
  geom_smooth(size=0.5)
ggplotly(p)
```

```{r}
## GGally - statistical view 
ggpairs(data = dataSet, columns = c("TotalWorkingYears","MonthlyIncome"), title = "TotalWorkingYears & MonthlyIncome", columnLabels = c("TotalWorkingYears", "MonthlyIncome($)"),ggplot2::aes(color = "Corr"))+
  theme(plot.title = element_text(hjust = 0.5, size=14, face="bold"))
```
